---
title: "Stat525 IE"
author: "Harry Lu 吕达"
date: "2017年2月24日"
output: html_document
---
#Step 1
###Read data, convert factor to numeric
```{r,warning=FALSE,echo=FALSE}
bird=read.table("http://www.stat.ucla.edu/projects/datasets/birds.dat",header=TRUE,sep=",",fill=TRUE)
bird$Elevation=as.numeric(as.character(bird$Elevation))
bird$Profile.Area=as.numeric(as.character(bird$Profile.Area))
bird$Height=as.numeric(as.character(bird$Height))
bird$Half.height=as.numeric(as.character(bird$Half.height))
bird$Latitude=as.numeric(as.character(bird$Latitude))
bird$Longitude=as.numeric(as.character(bird$Longitude))
bird$No..Species=as.numeric(as.character(bird$No..Species))
bird$Total.density=as.numeric(as.character(bird$Total.density))
bird

library(car)
```

#Questions
###(a) Are there more species of birds in denser forests? `No..Species` and `Profile.Area`
```{r,echo=FALSE}
scatterplot(No..Species~Profile.Area,data=bird)
```

###(b) Are birds concentrated more densely in areas where vegetation is more dense? `Total.density` and `Profile.Area`
```{r,echo=FALSE}
scatterplot(Total.density~Profile.Area,data=bird)
```

###(c) Is there more diversity in locations at higher latitudes, or lower latitudes? `No..Species` and `Latitude`
```{r,echo=FALSE}
scatterplot(No..Species~Latitude,data=bird)
```

###(d) Which technical measure of vegetation structure is the best predictor of the number of bird species in a forest. `???` and `No..Species`
```{r,echo=FALSE}
scatterplotMatrix(~SITE+Elevation+Profile.Area+Height+Half.height+Latitude+Longitude+No..Species+Total.density,data=bird)
```



#Modeling

#Plot scatterplot matrix
```{r,echo=FALSE}
scatterplotMatrix(~SITE+Elevation+Profile.Area+Height+Half.height+Latitude+Longitude+No..Species+Total.density,data=bird)
```

#Relationship between `Number of Species` and `Total Density`
```{r,echo=FALSE}
scatterplotMatrix(~No..Species+Total.density,data=bird)
```

#Analyse variables
###1. Remove `Latitude` and `Longitube`
```{r,echo=FALSE}
scatterplotMatrix(~Latitude+Longitude+No..Species+Total.density,data=bird)
```

###2. Relationship between `Profile Area`, `Height` and `Half.height`
```{r,echo=FALSE}
scatterplotMatrix(~Profile.Area+Height+Half.height,data=bird)
```

###3. Add a column `ln(Elevation)`
```{r,echo=FALSE}
bird$log_Elevation=log(bird$Elevation)
scatterplotMatrix(~Elevation+log_Elevation+No..Species+Total.density,data=bird)
```



#none use
```{r,echo=FALSE}
#plot(resid~Total.density)
#line1=lm(resid~Total.density)
#bird$resid <- bird$No..Species - summary(line1)$coef[1,1] - summary(line1)$coef[2,1]*Total.density
#bird$squareResid <- (bird$resid)^2
#mean(bird$resid)
#mean(bird$squareResid)

#plot(squareResid~Total.density)
#line2=lm(squareResid~Total.density)
#abline(line2)
#summary(line2)$coef
```

