---
output:
  pdf_document: default
  html_document: default
---
#Read in Data
```{r,warning=FALSE,echo=FALSE}
bird=read.table("http://www.stat.ucla.edu/projects/datasets/birds.dat",header=TRUE,sep=",",fill=TRUE)

# convert list rows into easier to parse format.
for(x in 2:9){
  bird[,x] = as.numeric(as.character(bird[,x]))
}

head(bird)
library(car)
```
###This data is from a study done on the number of breeding pairs of birds in 40 locations around the American Pacific coastline as well as closer inland. Taken during the spring of 1994, the sites were all around 5 hetacres in size and situated in relativly homogenous habitats. The number breeding pairs of bird species was counted by sight and by analyzing sound recordings. The variables are defined by the researchers as follows:

###SITE...................The site in which the data was taken

###Elevation...................Meters above sea level

###PROFILE AREA...................Roughly, a measure of the total amount of foliage at a site. Technically, this number is obtained by plotting vegetation density versus height above ground and taking the total area under the graph. The units, however, are not typical area units, as vegetation density is the reciprocal of a distance; the units are specially designed to measure profile area, and are called f.p. units (foliage profile units).

###HEIGHT...................Roughly, the height of the top of the forest canopy. Technically, the height above which horizontal distance to semi-obscurity exceeds 30 meters (a height from which line of sight is over 30 meters). Given in meters.  

###HALF-HEIGHT...................Roughly, the height that divides total foliage in two: half the forest is above it, and half below. Technically, the median of the graph of vegetation density versus height. Given in meters. 

###LATITUDE...................The latitude of the site. Given in degrees.

###LONGITUDE..................The longitude of the site. Given in degrees.

###NO. SPECIES................The number of different species of breeding bird pairs. 

###TOTAL DENSITY..............A measure of the bird population density. It is simply the number of pairs of birds per hectare.   
			   
###There is data missing from the Elevation, Profile.Area, Height, and Half.height variables. The missing points of elevation seem to be random however, the missing data points for Profile.Area, Height, and Half.height follow the following conditions: 
###A site missing its Half.height will also not have its Profile area and vice versa. 
###If a site is missing its Height variable, it will also be missing its Profile.Area and Half.height variables. 

###A likely explanation for this is that Profile.Area is related in some way to Half.height, according to the description of the variables, Half.height and Profile.Area are both calculated from vegetation density with respect to height. So it would also make sense that in the cases where Height was not calculatable that the data points for Profile.Area and Half.height are missing as well. We check our theory by plotting Profile.Area vs Height, Half.height vs Height, and Half.height vs Profile.Area.

\newpage

```{r, warning=FALSE, echo=FALSE}

ols_profile_vs_height <- lm(Profile.Area ~ Height, data=bird, na.action = na.exclude)

profile_vs_height <- ggplot(bird, aes(x = Height, y = Profile.Area))

profile_vs_height + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_profile_vs_height)



ols_half_vs_height <- lm(Half.height ~ Height, data=bird, na.action = na.exclude)

half_vs_height <- ggplot(bird, aes(x = Height, y = Half.height))

half_vs_height + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_half_vs_height)



ols_half_vs_profile <- lm(Half.height ~ Profile.Area, data=bird, na.action = na.exclude)

half_vs_profile <- ggplot(bird, aes(x = Profile.Area, y = Half.height))

half_vs_profile + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_half_vs_profile)
```

###Our evidence supports our claim that these three variables are correlated with eachother. In particular, we found small p values for the slopes of our estimates. This does not necessarily mean that the relationships must be linear however as just by eyeballing the first graph, we can see faintly a logarithmic spread among the points not including the leftmost one. We should next check the residual plots to make sure that the relationship is indeed linear now.

\newpage
# Residual Plots

```{r, warning=FALSE, echo=FALSE}
#Profile.Area vs Height residual graph
bird$pred_profile.Area_from_height <- predict(ols_profile_vs_height)

bird$resid_profile_vs_height <- resid(ols_profile_vs_height)

ols_resid_profile_vs_height_vs_profile <- lm(resid_profile_vs_height ~ pred_profile.Area_from_height, data = bird)

resid_profile_vs_height_vs_profile <- ggplot(bird, aes(x = pred_profile.Area_from_height, y = resid_profile_vs_height))

resid_profile_vs_height_vs_profile + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_resid_profile_vs_height_vs_profile)



#Half.height vs height residual graph
bird$pred_half.height_from_height <- predict(ols_half_vs_height)

bird$resid_half_vs_height <- resid(ols_half_vs_height)

ols_resid_half_vs_height_vs_half <- lm(resid_half_vs_height ~ pred_half.height_from_height, data = bird)

resid_half_vs_height_vs_half <- ggplot(bird, aes(x = Half.height, y = resid_half_vs_height))

resid_half_vs_height_vs_half + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_resid_half_vs_height_vs_half)
  


#Half.height vs Profile.Area residual graph 
bird$pred_half.height_from_profile.Area <- predict(ols_half_vs_profile)

bird$resid_half_vs_profile <- resid(ols_half_vs_profile)

ols_resid_half_vs_profile_vs_profile <- lm(resid_half_vs_profile ~ Profile.Area, data = bird)

resid_half_vs_profile_vs_profile <- ggplot(bird, aes(x = Profile.Area, y = resid_half_vs_profile))

resid_half_vs_profile_vs_profile + geom_point() + stat_smooth(method = "lm", col = "red", fullrange=TRUE)

summary(ols_resid_half_vs_profile_vs_profile)

```
#TODO: add analysis
###The p values of some residual plots are small which indicates that . The residuals of our estimates of Profile.Area bsed on Height has a small negative slope. 

#Questions
###These are the questions posed by researchers 
###(a) Are there more species of birds in denser forests? `No..Species` and `Profile.Area`
```{r,echo=FALSE}
scatterplot(No..Species~Profile.Area,data=bird)
```

###(b) Are birds concentrated more densely in areas where vegetation is more dense? `Total.density` and `Profile.Area`
```{r,echo=FALSE}
scatterplot(Total.density ~ Profile.Area, data=bird)
ols1 <- lm(Total.density ~ Profile.Area + Half.height, data = bird)
summary(ols1)
```
###We can see that along with Half.height, Profile.Area may be a good estimator of Total.density


###(c) Is there more diversity in locations at higher latitudes, or lower latitudes? `No..Species` and `Latitude`
```{r,echo=FALSE}
scatterplot(No..Species~Latitude,data=bird)
```

###(d) Which technical measure of vegetation structure is the best predictor of the number of bird species in a forest. `???` and `No..Species`
```{r,echo=FALSE}
scatterplotMatrix(~SITE+Elevation+Profile.Area+Height+Half.height+Latitude+Longitude+No..Species+Total.density,data=bird)
```

###(e) Other questions to consider:
###Is number of species correlated with the total density of birds?


#Modeling

#Plot scatterplot matrix
```{r,echo=FALSE}
scatterplotMatrix(~SITE+Elevation+Profile.Area+Height+Half.height+Latitude+Longitude+No..Species+Total.density,data=bird)
```

#Relationship between `Number of Species` and `Total Density`
```{r,echo=FALSE}
scatterplotMatrix(~No..Species+Total.density,data=bird)
```

#Analyse variables
###1. Remove `Latitude` and `Longitube`
```{r,echo=FALSE}
scatterplotMatrix(~Latitude+Longitude+No..Species+Total.density,data=bird)
```
###It seems that latitude and longitute are not well correlated with No..Species or Total.density as their plots look like null plots. This may be because the area that the samples were taken were geographically close to eachother. 


###2. Relationship between `Profile Area`, `Height` and `Half.height`
```{r,echo=FALSE}
scatterplotMatrix(~Profile.Area+Height+Half.height,data=bird)
```
###It seems Profile.Area, Height, and Half.height are all heavily correlated with eachother and so prehaps it is enough to include just one of the three as a parameter in a model that we choose.


###3. Add a column `ln(Elevation)`
```{r,echo=FALSE}
bird$log_Elevation=log(bird$Elevation)
scatterplotMatrix(~Elevation+log_Elevation+No..Species+Total.density,data=bird)
```



#none use
```{r,echo=FALSE}
#plot(resid~Total.density)
#line1=lm(resid~Total.density)
#bird$resid <- bird$No..Species - summary(line1)$coef[1,1] - summary(line1)$coef[2,1]*Total.density
#bird$squareResid <- (bird$resid)^2
#mean(bird$resid)
#mean(bird$squareResid)

#plot(squareResid~Total.density)
#line2=lm(squareResid~Total.density)
#abline(line2)
#summary(line2)$coef
```

